---
title: Plot Missing Values
author: ''
date: '2018-05-22'
slug: plot-missing-values
categories:
  - r tips
tags:
  - EDA
  - visualization
  - function
---


> In this post, we will explore **visdat** package to visualize missing values. Then, we will recreate a similar plot using **ggplot2** package. Next, using the **ggplot2** method, we will visualize other numeric or character values in a dataset. Finally, using the **ggplot2** method, we will create a custom function that will plot any value in any dataset.

There are many ways to inspect and find about missing values in a dataset. For example, `summary()` function from base R produces descriptive statistics and number of `NA`'s. Let's use that with **airquality** dataset, that has some missing values.

```{r message=FALSE, warning=FALSE}
summary(airquality)
```

From the output of `summary()` function, we can see columns/variables **Ozone** has 36 missing values (`NA`'s) and column **Solar.R** has 7 missing values. 

You can also inspect the whole dataset as in a spreadsheet view using `View()`.

```{r eval=FALSE, message=FALSE, warning=FALSE}
View(airquality)
```
