---
title: Collecting Tweets with R
author: Masood Sadat
date: '2018-06-14'
slug: collecting-tweets-with-r
categories:
  - r tips
tags:
  - twitter
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(twitteR)

api_key <- "Xt0oXgKSo6LdrDGoXaWr50CYw"
api_secret <- "rdjVhLuwYY9VKKU2NAIU5WGM8QKUaALZ6pMiA0tdhin6i3j5F4"
access_token <- "987749453288394753-cvmDm0t3MhOOEw4PFmDFRqjcvwnSrag"
access_secret <- "8G344YbekypsnXLgKNqGBc7Vx88b5JbLFxHcdf8AZOg5a"
```

In this post I am going to download tweets using **twitteR** package.

1. To being, you have to have a twitter account. If you do not have, create a twitter account on [twitter.com](https://twitter.com).

2. Create a twitter application on [apps.twitter.com](https://apps.twitter.com). There, you have to insert a number of information, see the following screen shot.


3. Once, you created an application, you will have access to _API Key_ and _API Secret_ in the **Keys and Access Tokens** tab. In the same tab, you can create _access tokens_ too. Store these information in R, but before that you have to do the step 4.

4. Go to RStudio: install and attach **twitteR** package.

```{r eval=FALSE}
install.packages("twitteR")
library(twitteR)
```

