---
title: Linear regression with R
author: Masood
date: '2018-06-12'
slug: linear-regression-with-r
categories:
  - r tips
tags:
  - statistics
---


# Estimating linear regression

Before starting, I am going to load a number of utility libraries for this session.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(plotly)
```

A simple linear regression using **mtcars** dataset. I want to look at how weight of a car is associated with how many miles a car travels per gallon of gasoline--an indicator of fuel economy. I am going to look at their summary and scatter plot. For scatter plot, I use base-R's `plot()` function.


```{r message=FALSE, warning=FALSE}
summary(mtcars[, c("wt", "mpg")])
plot(mtcars$wt, mtcars$mpg)
```

It seems like heavier cars have lower mileage Next, I regress mpg on weight using linear regression.

```{r message=FALSE, warning=FALSE}
lm(mpg ~ wt, data = mtcars)
```

The output from `lm()` function quantifies a negative relationship between weight and mileage. The output however does not provide other useful information such as confidence interval and r-squared, which can be accessed if we feed the `lm` function into `summary()` function.

```{r message=FALSE, warning=FALSE}
summary(lm(mpg ~ wt, data = mtcars))
```
