---
title: Explore Correlation Among Variables with Corrplot
author: Masood Sadat
date: '2018-06-14'
slug: explore-correlation-among-variables-with-corrplot
categories:
  - r tips
tags:
  - EDA
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(corrplot)
```

Exploring relationship among variables in a data set is never easier with **corrplot** package and function of the same name. **corrplot** takes in a correlation matrix and returns a visual representation of that correlation matrix.

Try it for yourself. 

1. Install and load **corrplot** package

```{r eval=FALSE}
install.packages("corrplot")
library(corrplot)
```

2. Estimate correlation between all variables of **mtcars** dataset using `cor`, and then plot the correlations using `corrplot()`. You have to specify a method, and there are quite a few of them--here I demonstrate two of them: _square_ and _number_

```{r}
mat <- cor(mtcars)
corrplot(mat, method = "square")
corrplot(mat, method = "number")
```

To explore more, see the help (`?corrplot`) and visit the Introduction to corrplot Package post on [CRAN](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html).
