---
title: Hypothesis testing with R
author: Masood Sadat
date: '2018-06-14'
slug: hypothesis-testing-with-r
categories:
  - r tips
tags:
  - statistics
---

## A Meaningful Relationship

There seems to be a relatinship between weight (`wt`) and mileage (`mpg`) from the **mtcars** based on observing them on a scatter plot--perhaps a negative one.

```{r message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) + geom_point()
```

Fit a linear model to estimate the relationship between weight and mpg, and tidy the output. Looking at the statistics and p-value, I can reject the null hypothesis (no relationship).

```{r message=FALSE, warning=FALSE}
library(broom)
library(dplyr)
tidy(lm(mpg ~ wt, data = mtcars))
```

Normally, the above estimation would suffice to answer a hypothesis question. For education, let's create a null (no relationship) distribution of slopes using **infer** package.

```{r message=FALSE, warning=FALSE}
# install.packages("infer")
library(infer)
perm_slopes <- mtcars %>%
  specify(formula = mpg ~ wt) %>% # specify the linear model
  hypothesize(null = "independence") %>% # provide null hypothesis that two variables are independent
  generate(reps = 1000, type = "permute") %>% # describe how to generate the sampling distribution
  calculate(stat = "slope") # calculate the statistic of interest

obs_slope <- tidy(lm(mpg ~ wt, data = mtcars)) %>% .[2,2] # observed slope
```

Then, plot the permuted slopes using a histogram and add the observed slope as a vertical line. I can see the observed slope does not overlap even one of the 1000 permuted slopes, which is a visual way of finding that probability of observing null hypothesis is less than 0.1%. This means, I can reject the null hypothesis.

```{r message=FALSE, warning=FALSE}
ggplot(perm_slopes, aes(stat)) + geom_histogram() + # histogram distribution of slope of null hypothesis
  geom_vline(xintercept = obs_slope, color = "red") # add a vertical line for observed slope
```

## A Not-Meaningful Relationship

To contrast, let's look at the relationship between real axle ratio (`drat`) and "1/4 mile time" (`qsec`) from the same **mtcars** data, which appears to have no relationship based on observing them on a scatter plot.

```{r}
ggplot(mtcars, aes(drat, qsec)) + geom_point()
```

Fitting a linear model to estimate the linear relationship between `drat` and `qsec`, I cannot reject the null hypothesis that there is a meaningful relationship between the two variables. Note the statistics and p-value values.

```{r}
tidy(lm(qsec ~ drat, data = mtcars))
obs_slope <- tidy(lm(qsec ~ drat, data = mtcars)) %>% .[2,2]
```

Let's estimate the null (no relationship) distribution using the **infer** package.

```{r}
perm_slopes <- mtcars %>%
  specify(qsec ~ drat) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "slope")
```

Next, I am going to use the permuted slops to plot a histogram, and add a red vertical line that represents the observed slope of relationship between `qsec` and `drat`. I can see that the observed slope overlaps the permuted slopes (null distribution), and once again I come to the same conclusionthat there is no meaningful relationship between `qsec` and `drat`.

```{r message=FALSE, warning=FALSE}
# ggplot(perm_slopes, aes(stat)) + geom_histogram() +
#   geom_vline(xintercept = obs_slope, color = "red")
```
